---
title: 'Market Trends' 
author: Stephanie Langeland
always_allow_html: yes
output:
  html_document:
    keep_md: true
    toc: true
    number_sections: true
---

# File Info

Date: 04-18-2017

Author: Stephanie Langeland

File Name: `10_market_trends_20180418_v1.Rmd`

Version: 78

Previous Version/File: `10_market_trends_20180411_v1.Rmd`

Dependencies: None

Purpose:

  1) Explore open vs. close data and intraday data on FOMC and non-FOMC dates.
  
  2) Decide whether to use open vs. close and/or intraday data as response 
  variables in the final LDA model.

Input File(s):

  * `open_close_data.rds`

  * `cla_intraday.rds`

  * `ct10govt_intraday.rds`
  
  * `spx_intraday.rds`
  
  * `ussoc_intraday.rds`
  
  * `vix_intraday.rds`
  
Output File(s): `/code/market_trends_images/`

Data Output: None

Required by: QMSS Master's Thesis

Status: Complete

Machine: 2018 MacBook Pro

R version: `r {c( R.Version()$version.string, R.Version()$nickname)}`

```{r Setup, include = FALSE, results = 'hide', warning = FALSE}
#install.packages('knitr')
library(knitr)
opts_chunk$set(fig.path = '10_market_trends_images/',
               cache.path = 'cache/',
               cache = FALSE,
               echo = TRUE,
               message = FALSE,
               warning = FALSE
)
```

# Open and Close Data

```{r, results = "hide"}
## :::::::::::::::::::: Import Open and Close Data:
options(scipen = 999)

oc_data <- readRDS("/Users/stephanielangeland/Desktop/Columbia/masters_thesis/Langeland_thesis/input_output_files/open_close_data.rds")

str(oc_data)

##View(oc_data)

## :::::::::::::::::::: Add Recession Dates:
oc_data$recession_date <- ifelse(
  oc_data$release_date > "2001-03-01" &
    oc_data$release_date < "2001-11-01" ,
  Inf,
  0
)

oc_data$recession_date <- ifelse(
  oc_data$release_date > "2007-12-01" &
    oc_data$release_date < "2009-06-01" ,
  Inf,
  oc_data$recession_date
)



## :::::::::::::::::::: Create subsets of the dataset - FOMC release dates vs. non-release dates
library(dplyr)

releases_sub <- oc_data %>%
  filter(
    !is.na(stmt_text) |
    !is.na(min_text)
  )

no_releases_sub <- oc_data %>%
  filter(
    is.na(stmt_text) &
    is.na(min_text)
  )
```

## U.S. corporate credit spread (BICLB10Y) - last price only

```{r "U.S. Corporate Credit Spread on FOMC vs. Non-FOMC Dates", warning = FALSE, message = FALSE}
## :::::::::::::::::::: Plot:
library(ggplot2)

spread_plot <- ggplot() +
  geom_area(aes(x = oc_data$release_date,
                y = oc_data$recession_date,
                alpha = 0.05),
            fill = "gray83",
            show.legend = F) +
  geom_point(aes(x = no_releases_sub$release_date,
                y = no_releases_sub$BICLB10Y_LAST_PRICE,
                color = "Non-FOMC Dates"
                )) +
  geom_point(aes(x = no_releases_sub$release_date,
                y = no_releases_sub$BICLB10Y_LAST_PRICE,
                color = "Non-FOMC Dates",
                text = paste("Non-FOMC Date: ",
                                as.character(no_releases_sub$release_date),
                                "\n",
                                "U.S. corporate credit spread (last price): ",
                                no_releases_sub$BICLB10Y_LAST_PRICE)),
             size = 0.9) +
  geom_point(aes(x = releases_sub$release_date,
                y = releases_sub$BICLB10Y_LAST_PRICE,
                color = "FOMC Dates",
                text = paste("FOMC Meeting Materials Release Date: ",
                                as.character(releases_sub$release_date),
                                "\n",
                                "U.S. corporate credit spread (last price): ",
                                releases_sub$BICLB10Y_LAST_PRICE))) +
  labs(title = "U.S. Corporate Credit Spread on FOMC vs. Non-FOMC Dates",
       x = "Date",
       y = "U.S. Corporate Credit Spread") +
  scale_colour_manual("Legend", 
                      breaks = c("FOMC Dates", 
                                 "Non-FOMC Dates"),
                      values = c("magenta1", "gray45")) +
  theme(
    panel.border = element_rect(fill = NA),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "gray",
                                    size = 0.25),
    panel.grid.minor = element_line(colour = "gray",
                                    size = 0.25)
  ) 

spread_plot
```

__Interactive Plot:__

```{r fig.height = 7, fig.width = 15, warning = FALSE, message = FALSE}
library(plotly)

ggplotly(spread_plot,
         tooltip = c("text"))
```

```{r}
## :::::::::::::::::::: Tables:
mean(releases_sub$BICLB10Y_LAST_PRICE, 
     na.rm = T) ## mean on FOMC dates

mean(no_releases_sub$BICLB10Y_LAST_PRICE, 
     na.rm = T) ## mean on non-FOMC dates
```

Notes: Average U.S. corporate credit spread is higher on release days.

## 10-year constant maturity rate (H15T10Y)

```{r, message = FALSE, warning = FALSE}
library(QMSS)

Tab(releases_sub$H15T10Y_change_cat)  ## FOMC dates

Tab(no_releases_sub$H15T10Y_change_cat) ## non-FOMC dates
```

Notes: Doesn't seem like FOMC releases change the trend at all. 

## S&P 500 Index (SPX Index)

```{r}
Tab(releases_sub$SPX_change_cat)  ## FOMC dates

Tab(no_releases_sub$SPX_change_cat) ## non-FOMC dates
```

Notes: Doesn't seem like FOMC releases change the trend at all. 

## 3-month U.S. overnight index swap rate (USSOC Currency)

```{r}
Tab(releases_sub$USSOC_change_cat)  ## FOMC dates

Tab(no_releases_sub$USSOC_change_cat) ## non-FOMC dates
```

Notes: Doesn't seem like FOMC releases change the trend at all. 

## CBOE VIX Index (VIX) 

```{r}
Tab(releases_sub$VIX_change_cat)  ## FOMC dates

Tab(no_releases_sub$VIX_change_cat) ## non-FOMC dates
```

Notes: Doesn't seem like FOMC releases change the trend at all. 

## Crude oil price (CLA Comdty)

```{r}
Tab(releases_sub$CLAComdty_change_cat)  ## FOMC dates

Tab(no_releases_sub$CLAComdty_change_cat) ## non-FOMC dates
```

Notes: 

  * Never 0 (no volatility in the 3-month U.S. overnight index swap rate) on FOMC dates
  
  * 3-month U.S. overnight index swap rate usually goes down on FOMC dates so the fed might calm markets - could
  indicate successful transparency because they don't disrupt the market too much.  The opposite
  is true overall on non-FOMC dates.

## Moody’s BAA corporate bond yield (MOODCBAA Index) - last price only 

```{r}
mean(releases_sub$MOODCBAA_LAST_PRICE, na.rm = T)  ## FOMC dates

mean(no_releases_sub$MOODCBAA_LAST_PRICE, na.rm = T) ## non-FOMC dates
```

Notes: Doesn't seem like FOMC releases change the trend at all.

## CBOE Interest Rate 10-year Treasury Note (^TNX)

```{r}
Tab(releases_sub$TNX_change_cat)  ## FOMC dates

Tab(no_releases_sub$TNX_change_cat) ## non-FOMC dates
```

Notes: T-bill rate usually goes up on FOMC dates but it usually go down overall
on non-FOMC dates.  Historically, they've lowered rates more than they've lowered
them during the sample period:
    
```{r}
Tab(oc_data$TNX_change_cat)
```
    
# Intraday Data 

## S&P 500 Index (SPX Index)

```{r, results = "hide"}
spx <- readRDS("/Users/stephanielangeland/Desktop/Columbia/masters_thesis/Langeland_thesis/input_output_files/spx_intraday.rds")

str(spx)

releases_spx_sub <- spx %>%
  filter(
    !is.na(type) |
    !is.na(type)
  )

releases_spx_sub <- na.omit(releases_spx_sub)

no_releases_spx_sub <- spx %>%
  filter(
    is.na(type) &
    is.na(type)
  )
```

```{r}
spx_rel <- colMeans(releases_spx_sub[, 5:62],
                     na.rm = T) ## fomc dates

#spx_rel

spx_norel <- colMeans(no_releases_spx_sub[, 5:62],
                       na.rm = T) ## non-fomc dates

#spx_norel 
```

```{r "Intraday SPX"}
plot(spx_rel,
     type = "p",
     col = "red",
     ylab = "")

par(new = TRUE)

plot(spx_norel,
     type = "l",
     col = "black")
```

Notes: Doesn't look like there's a difference in the 2 trends.

## 3-month U.S. overnight index swap rate (USSOC Currency)

```{r, results = "hide"}
ussoc <- readRDS("/Users/stephanielangeland/Desktop/Columbia/masters_thesis/Langeland_thesis/input_output_files/ussoc_intraday.rds")

str(ussoc)

#View(ussoc)

releases_ussoc_sub <- ussoc %>%
  filter(
    !is.na(type) |
    !is.na(type)
  )

releases_ussoc_sub <- na.omit(releases_ussoc_sub)

##View(releases_ussoc_sub)

no_releases_ussoc_sub <- ussoc %>%
  filter(
    is.na(type) &
    is.na(type)
  )

#View(no_releases_ussoc_sub)
```

```{r}
ussoc_rel <- colMeans(releases_ussoc_sub[, 5:90],
                       na.rm = T) ## fomc dates

#ussoc_rel

ussoc_norel <-colMeans(no_releases_ussoc_sub[, 5:90],
                       na.rm = T) ## non-fomc dates

#ussoc_norel
```

```{r "Intraday USSOC"}
plot(ussoc_rel,
     type = "p",
     col = "red",
     ylab = "")

par(new = TRUE)

plot(ussoc_norel,
     type = "l",
     col = "black")
```

Notes: A little bit of a difference in trend.

## CBOE VIX Index (VIX) 

```{r, results = "hide"}
vix <- readRDS("/Users/stephanielangeland/Desktop/Columbia/masters_thesis/Langeland_thesis/input_output_files/vix_intraday.rds")

str(vix)

#View(vix)

releases_vix_sub <- vix %>%
  filter(
    !is.na(type) |
    !is.na(type)
  )

releases_vix_sub <- na.omit(releases_vix_sub)

#View(releases_vix_sub)

no_releases_vix_sub <- vix %>%
  filter(
    is.na(type) &
    is.na(type)
  )

#View(no_releases_vix_sub)
```

```{r}
vix_rel <- colMeans(releases_vix_sub[, 5:34],
                       na.rm = T) ## fomc dates

#vix_rel

vix_norel <- colMeans(no_releases_vix_sub[, 5:34],
                       na.rm = T) ## non-fomc dates

#vix_norel
```

```{r "Intraday VIX"}
plot(vix_rel,
     type = "p",
     col = "red",
     ylab = "")

par(new = TRUE)

plot(vix_norel,
     type = "l",
     col = "black")
```

Notes: Little bit of a difference in trend - release dates are trending lower.

## Crude oil price (CLA Comdty)

```{r, results = "hide"}
cla <- readRDS("/Users/stephanielangeland/Desktop/Columbia/masters_thesis/Langeland_thesis/input_output_files/cla_intraday.rds")

str(cla)

#View(cla)

releases_cla_sub <- cla %>%
  filter(
    !is.na(type) |
    !is.na(type)
  )

releases_cla_sub <- na.omit(releases_cla_sub)

#View(releases_cla_sub)

no_releases_cla_sub <- cla %>%
  filter(
    is.na(type) &
    is.na(type)
  )

#View(no_releases_cla_sub)
```

```{r}
cla_rel <- colMeans(releases_cla_sub[, 5:41],
                       na.rm = T) ## fomc dates

#cla_rel

cla_norel <- colMeans(no_releases_cla_sub[, 5:41],
                       na.rm = T) ## non-fomc dates

#cla_norel
```

```{r "Intraday CLA Comodty"}
plot(cla_rel,
     type = "p",
     col = "red",
     ylab = "")

par(new = TRUE)

plot(cla_norel,
     type = "l",
     col = "black")
```

Notes: Little bit of a difference in trend - release dates are trending higher.

## 10-year Treasury yield (CT10 Govt)

```{r, results = "hide"}
ct10 <- readRDS("/Users/stephanielangeland/Desktop/Columbia/masters_thesis/Langeland_thesis/input_output_files/ct10govt_intraday.rds")

str(ct10)

#View(ct10)

releases_ct10_sub <- ct10 %>%
  filter(
    !is.na(type) |
    !is.na(type)
  )

releases_ct10_sub <- na.omit(releases_ct10_sub)

#View(releases_ct10_sub)

no_releases_ct10_sub <- ct10 %>%
  filter(
    is.na(type) &
    is.na(type)
  )

#View(no_releases_ct10_sub)
```

```{r}
ct10_rel <- colMeans(releases_ct10_sub[, 5:90],
                       na.rm = T) ## fomc dates

#ct10_rel

ct10_norel <- colMeans(no_releases_ct10_sub[, 5:90],
                       na.rm = T) ## non-fomc dates

#ct10_norel
```

```{r "Intraday CT10 GOVT"}
plot(ct10_rel,
     type = "p",
     col = "red",
     ylab = "")

par(new = TRUE)

plot(ct10_norel,
     type = "l",
     col = "black")
```

Notes: Doesn't look like there's a difference in the 2 trends.

# Conclusion: Only use open vs. close data in the final LDA model.
